---
title: "Cleaning Data in R"
author: "Bourbon0212"
date: "2018年7月14日"
output: html_document
---
<h2 id = '0'></h2>
* [Exploring Raw Data](#1)    
* [Tidying Data](#2)    
* [Preparing Data for Analysis](#3)    
* [Case Study](#4)

<h2 id = '1'></h2>
# **Exploring Raw Data**    

**Functions Helping Understanding Raw Data**    

* **class()**: Class of data object.    
* **dim()**: Dimensions of data.    
* **names()**: Cloumn names.    
* **str()**: Preview of data with helpful details.    
* **glimpse()**: Better version of str() from ```dplyr```
* **summary()**: Summary of data.     
* **head(,```row```)**: Allow you to quickly see the top a dataset. The default of ```row``` = 6.   
* **tail(,```row```)**: Allow you to quickly see the bottom a dataset. The default of ```row``` = 6.    
* **hist(```dataframe$column```)**: Visualize data with histogram.
* **plot(```dataframe$column```, ```dataframe$column```)**: Visualize data as a series of (x, y) coordinates on a two-dimensional plane with scatter plot.

```{r message = F}
bmi <- read.csv('https://assets.datacamp.com/production/repositories/34/datasets/a0a569ebbb34500d11979eba95360125127e6434/bmi_clean.csv')
#Getting a feel for your data
class(bmi)
dim(bmi)
names(bmi)
#Viewing the structure of your data.
str(bmi)
library(dplyr)
glimpse(bmi)
summary(bmi)
#Histogram of BMIs from 2008
hist(bmi$Y2008)
#Scatter plot comparing BMIs from 1980 to those from 2008
plot(bmi$Y1980, bmi$Y2008)
```

<h2 id = '2'></h2>
# **Tidying Data**

**Principles of Tidy Data**   

* Observation as rows.    
* Variables as columns.   
* One type of observational unit per table.
Otherwise, it's a messy data.

**Common Symptoms of Messy Data**   

* Column headers are values, not variable names. -> gather()    
* Variables are stored in both rows & columns. -> spread()    
* Multiple variables are stored in one column. -> separate()    
* A single observational unit is stored in multiple tables.   
* Multiple types of observational units are stored in the same table.

**Tidyr: Gather & Spread**    

* **gather(```wide_df, my_key, my_val, -col```)**: Gather columns to key-value pairs. Makes wide datasets long.    
* **spread(```long_df, my_key, my_val```)**: Spread key-value pairs to columns. Makes long datasets wide.    

```{r}
library(tidyr)
bmi_long <- gather(bmi, year, bmi_val, -Country)
head(bmi_long)
bmi_wide <- spread(bmi_long, year, bmi_val)
head(bmi_wide)
```
Compare bmi, bmi_long & bmi_wide, notice that bmi_wide and bmi are indeed the same.     

**Tidyr: Separate & Unite**   

* **separate(```data, col, into(c("col1", "col2")), sep = '*'```)**: Separate one column into multiple.    
* **unite(```data, col, ...(bare names of columns), sep = '*'```)**: Unite multiple columns into one.   

<h2 id = '3'></h2>
# **Preparing Data for Analysis**

**Types Conversion**    

* **class()**: Class of data object.    
* **as.```class```()**: Change the class of a object.   
* **library(lubridate)**: A package deal with date & time.```y```, ```m```, ```d```, ```h```, ```m```, ```s```, which stand for year, month, day, hour, minute, and second, respectively.   

```{r message = F}
library(lubridate)
students <- read.csv('https://assets.datacamp.com/production/repositories/34/datasets/f75a87dbbdf2cf79e2286f97b2af22146cb717b1/students_with_dates.csv')
str(students$dob)
str(students$Grades)
students$Grades <- as.character(students$Grades)
students$dob <- ymd(students$dob)
str(students$dob)
str(students$Grades)
```

**String Manipulation**    

* ```stringr```:    

    + **str_trim()**: Triming leading & trailing white space.   
    + **str_pad()**: Pad string with additional characters.   
    + **str_detect(```input```,```str```)**: Detect a pattern.    
    + **str_replace(```input```,```replaced str```,```new str```)**: Find & replace a pattern.
    + Reviews: **[Grepl & Grep](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#2.5)** in ```base``` for detecting string.   
    + Reviews: **[Sub & gsub](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#2.5)** in ```base``` for repalcing string.   
* ```base```:   
    + **tolower()**: Make all lowercase.    
    + **toupper()**: Make all uppercase.    
    
```{r message = F}
library(stringr)
str_trim(c("   Filip ", "Nick  ", " Jonathan"))
str_pad(c("23485W", "8823453Q", "994Z"), width = 9, side = "left", pad = "0")
head(students$dob)
head(str_detect(students$dob, '1997'))
str(students$sex)
students$sex <- str_replace(students$sex, 'F', 'Female')
students$sex <- str_replace(students$sex, 'M', 'Male')
str(students$sex)
```

**Missing & Special Values**    

* **is.na()**: If missing values are properly coded as ```NA```, this function will help you find them.   
* **any(is.na())** Like asking the question "Are there any NA values in my dataset?".   
* **complete.cases()**: See which rows have no missing values.    
* **na.omit()**: Remove all rows with any missing values.

**Outliers & Obvious Errors**   

A simple **histogram** or **boxplot**, displaying the distribution of a variable's values across all the observations can be key to identifying potential outliers as early as possible.    
```{r}
hist(students$absences)
boxplot(students$absences)
```

<h2 id = '4'></h2>
# **Case Study**

* Understanding the structure of your data.   
* Looking at your data.   
* Tidy your data.
* Finally, visualizing your data.   

So, let's start with understanding the structure of your data.  
```{r}
weather <- readRDS('D:/Downloads/weather.rds')
#Verify that weather is a data.frame.
class(weather)
#Check the dimensions.
dim(weather)
#View the column names.
names(weather)
```

Then, we're going to looking at your data.    
```{r}
#View the structure of the data.
str(weather)
#Look at the structure using dplyr's glimpse().
library(dplyr)
glimpse(weather)
#View a summary of the data.
summary(weather)
#Take a closer look of your data.
head(weather)
```
Well done! Not surprisingly, this dataset is pretty messy…for now :)    
It's time to tidy it up!    
```{r}
# Gather the columns
library(tidyr)
weather2 <- gather(weather, day, value, X1:X31, na.rm = TRUE)#na.rm決定是否排除NA
head(weather2)
#First remove column of row names
weather2 <- weather2[, -1]#'-1' means to remove col1.
#Spread the data
weather3 <- spread(weather2, measure, value)
#View the head
head(weather3)
#Remove X's from day column
weather3$day <- str_replace(weather3$day, 'X', '')
#Unite the year, month, and day columns
weather4 <- unite(weather3, date, year, month, day, sep = "-")
#Convert date column to proper date format using lubridates's ymd()
weather4$date <- ymd(weather4$date)
#Rearrange columns using dplyr's select()
weather5 <- select(weather4, date, Events, CloudCover:WindDirDegrees)
#View the head of 
head(weather5)
```
And now, it's time for strange & missing values.
```{r}
#Search for strange values in PrecipitationIn
## View the structure of weather5
str(weather5)
#Examine the first 20 rows of weather5. Are most of the characters numeric?
head(weather5, 20)
weather5$PrecipitationIn <- str_replace(weather5$PrecipitationIn, 'T', '0')
#Or using sub(): 'weather5$PrecipitationIn <- sub('T', 0, weather5$PrecipitationIn)'
#Convert characters to numerics
weather6 <- mutate_at(weather5, vars(CloudCover:WindDirDegrees), funs(as.numeric))
#check the dataset
str(weather6)
summary(weather6)
```
Look at the summary, besides strange values in PrecipitationIn, there are also proplem in Max.Gust.SpeedMPH, Max.Humidity & Mean.VisibilityMiles.
```{r}
#Find indices of NAs in Max.Gust.SpeedMPH
#Count missing values
sum(is.na(weather6))
ind <- which(is.na(weather6$Max.Gust.SpeedMPH))
#Look at the full rows for records missing Max.Gust.SpeedMPH
weather6[ind, ]
#It's unclear why these values are missing and there doesn't appear to be any obvious pattern to their missingness, so we'll leave them alone for now. 

#Search for extreme values in Max.Humidity
ind <- which(weather6$Max.Humidity >= 1000)
#Look at the data for that day
weather6[ind, ]
#Change 1000 to 100
weather6$Max.Humidity[ind] <- 100

#Look at summary of Mean.VisibilityMiles
summary(weather6$Mean.VisibilityMiles)
#Get index of row with -1 value
ind <- which(weather6$Mean.VisibilityMiles == -1)
#Look at full row
weather6[ind, ]
#Set Mean.VisibilityMiles to the appropriate value
weather6$Mean.VisibilityMiles[ind] <- 10
```
Finally, let's have a final look of our data.
```{r}
#Replace empty cells in events column
weather6$Events[weather6$Events == ""] = 'None'
#Print the first 6 rows of weather6
head(weather6)
```
