---
title: "Data Manipulation in R with dplyr"
author: "Bourbon0212"
date: "2018年7月16日"
output: html_document
---

* [Introduction to dplyr & tbls](#1)    
* [Select & Mutate](#2)    
* [Filter & Arrange](#3)    
* [Summarise & the Pipe Operator](#4)
* [Group_by & Working with Databases](#5)

<h2 id = '1'></h2>
# **Introduction to dplyr & tbls**
Before we start the section, we should load the package```dplyr``` which includes all the useful functions we need now.   
A tbl (pronounced tibble) is just a special kind of data.frame. They make your data easier to look at, but also easier to work with. On top of this, it is straightforward to derive a tbl from a data.frame structure using ```tbl_df()```.   
Notice:  dplyr verbs **don't** adapt the data unless you explicitly assign the result to a new variable except ```summarise()```.    
In ```dplyr```, we have five useful functions to manipulate data:         

* **Variables**   

    + **select()**: Removes columns from a dataset.   
    + **mutate()**: Uses data to build new columns of values.   
* **Observations**    

    + **filter()**: Removes rows from a dataset.   
    + **arrange()**: Reorders rows in a dataset.    
* **Groups**    

    + **summarize()**: Calculates summary statistics.    
    
* Reviews: **[Data Wrangling](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.1)** in **[Introduction to the Tidyverse](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3)**   
```{r message = F}
library(dplyr)
library(hflights)
hflights <- tbl_df(hflights)
#Take a look at hflights.
hflights
#Using a LUT(look up table) change labels of hflights.
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")
hflights$Carrier <- lut[hflights$UniqueCarrier]
lut <- c("A" = "carrier", "B" = "weather", "C" = "FFA", "D" = "security", "E" = "not cancelled")
hflights$Code <- lut[hflights$CancellationCode]
#Check your results by glimpsing at them.
glimpse(hflights)
```

<h2 id = '2'></h2>
# **Select & Mutate**     

## Select & its Helper Functions    

* **select(```df```,```cols to select```,```...```)**: Removes columns from a dataset.   

    + Use ```:``` to select a range of variables and ```-``` to exclude some variables.      
    + Use both variable's names as well as integer indexes.   
    + ```select()``` does not change the data frame it is called on; you have to explicitly assign the result of ```select()``` to a variable to store the result.   

* **Helper functions for variable selection**   
Pay attention here: When you refer to columns directly inside ```select()```, you don't use quotes.
    + **starts_with("X")**: every name that starts with "X".    
    + **ends_with("X")**: every name that ends with "X".    
    + **contains("X")**: every name that contains "X".    
    + **matches("X")**: every name that matches "X", where "X" can be a regular expression.   
    + **num_range("x", 1:5)**: the variables named ```x01```, ```x02```, ```x03```, ```x04``` and ```x05```.    
    + **one_of(x)**: every name that appears in x, which should be a character vector.    
    
```{r}
#Using ':' & '-' to make a consice selection.
select(hflights,Year:Diverted, - (DepTime:AirTime))
#Using helper function for a consice selection.
select(hflights, contains('Time'), contains('Delay'))
#Select four columns of hflights related to delay for data analysis later.
select(hflights, 10:13)
```

## Mutate   
* **mutate(```df```,```new col name = expressions```)**: Uses data to build new columns of values.    
```{r}
#Add multiple variables using mutate
m2 <- mutate(hflights, TotalTaxi = TaxiIn + TaxiOut, ActualGroundTime = ActualElapsedTime - AirTime, Diff = TotalTaxi - ActualGroundTime)
m2
```

<h2 id = '3'></h2>
# **Filter & Arrange**    

## Filter, Logical & Boolean Operator   

* **filter(```df```, ```logical test```)**: Filters out rows from a dataset based on their contents.    
    + Besides common logical operatoers, ```x %in% c(a, b, c)```, ```TRUE``` if ```x``` is in the vector ```c(a, b, c)```   
    + Using Boolean Operators: ```&```, ```|``` and ```!```, for 'and', 'or' and 'not' respectively.    
    
```{r}
#All flights where taxiing took longer than flying.
filter(hflights, (TaxiIn + TaxiOut) > AirTime)
#All flights flown by one of JetBlue, Southwest, or Delta.
filter(hflights, UniqueCarrier %in% c("JetBlue", "Southwest", "Delta"))
#All flights that were cancelled after being delayed.
filter(hflights, Cancelled == 1 & DepDelay > 0)
```

## Arrange    

* **Arrange(```df```,```col name to arrange by```,```col name of the next tie breaker```)**: Reorders rows in a dataset.    
    + By default, ```arrange()``` arranges the rows from smallest to largest.     
    + Use ```desc()``` to reverse this behavior.    
```{r}
#Arrange according to carrier and decreasing departure delays
arrange(hflights, UniqueCarrier, desc(DepDelay))
```

<h2 id = '4'></h2>
# **Summarise & the Pipe Operator**  

##Summarise & dplyr Aggregate Functions

* **summarise(```df```, ```new col name = expression```)**: Use your data to create a new dataset of summary statistics that describe the data.   
* Reviews: **[summarize()](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.3)** seems that there isn't much difference between ```summarise()``` & ```summarize()```
* **dplyr Aggregate Functions**
    + **first(x)**: The first element of vector ```x```.
    + **last(x)**: The last element of vector ```x```.
    + **nth(x, n)**: The nth element of vector ```x```.
    + **n()**: The number of rows in the data.frame or group of observations that ```summarise()``` describes.
    + **n_distinct(x)**: The number of unique values in vector ```x```.

```{r}
#Generate summarizing statistics for hflights
summarise(hflights,
          n_obs = n(),
          n_carrier = n_distinct(hflights$UniqueCarrier),
          n_dest = n_distinct(hflights$Dest))
```

## The Pipe Operator: Chaining your functions

* **object %>% function(____, arg2, arg3, ...) -> object %>% function(arg2, arg3, ...)**
```{r}
#Chain together mutate(), filter() and summarise()
hflights %>%
  mutate(RealTime = ActualElapsedTime + 100, mph = Distance / RealTime * 60) %>%
  filter(!is.na(mph), mph < 70) %>%
  summarize(n_less = n(),
            n_dest = n_distinct(Dest),
            min_dist = min(Distance),
            max_dist = max(Distance))
```

<h2 id = '5'></h2>
# **Group_by & Working with Databases**     

## Group_by
* **group_by(```df```,```col to group by```)**: Define groups within your data set. when calling ```summarise()```, summarising statistics are calculated for the different groups separately.    
* **rank()**: Takes a group of values and calculates the rank of each value within the group.    

**Question 1**    
How many airplanes only flew to one destination?    
```{r}
hflights %>%
  group_by(TailNum) %>% #TaulNum是班機編號
  summarise(ndest = n_distinct(Dest)) %>%
  filter(ndest == 1) %>%
  summarise(nplanes = n())
```
**Question 2**    
Find the most visited destination for each carrier.
```{r}
hflights %>%
  group_by(UniqueCarrier, Dest) %>%
  summarise(n = n()) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
```
## Working with Databases using dplyr

* **tbl_df()**: Change dataframe to tbl.    
* **tbl_dt()**: Change data table to tbl.    
* **tbl()**: Change database to tbl.    

```{r message = F, warning=FALSE}
library(dbplyr)
library(RMySQL)
library(DBI)
#Set up a connection to the mysql database
my_db <- src_mysql(dbname = "dplyr", 
                   host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                   port = 3306, 
                   user = "student",
                   password = "datacamp")

#Reference a table within that source: nycflights
nycflights <- tbl(my_db, "dplyr")

#glimpse at nycflights
glimpse(nycflights)

#Ordered, grouped summary of nycflights
nycflights %>%
  group_by(carrier) %>%
  summarize(n_flights = n(),
            avg_delay = mean(arr_delay)) %>%
  arrange(avg_delay)%>%
  head()
```