---
title: "Exploratory Data Analysis"
author: "Bourbon0212"
date: "2018年7月13日"
output: html_document
---
<h2 id = '0'></h2>
* [Categorical Data](#1)    
* [Numerical Data](#2)    
* [Numerical Summaries](#3)    
* [Case Study](#4)

<h2 id = '1'></h2>
# **Categorical Data**    

**Contingency Table**   

* Reviews: **[Factors](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#1.5)**   
* Reviews: **[Dataframe](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#1.6)**   
* levels(): To show the levels of the factors.   
* Use '$' to select a specific row or column.   
* table(): To generate 'Contingency table'.   
```{r message = F}
library(readr)
comics <- read_csv('https://assets.datacamp.com/production/course_1796/datasets/comics.csv')
comics[1,]
levels(comics$align)
table(comics$align, comics$gender)
```
**Dropping Levels**   

The contingency table from the last exercise revealed that there are some levels that have very low counts.   

* Reviews: **[filter()](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.1)** for **Data Wrangling**    
* droplevels(): Drops levels that don't exist in a dataset.   

Notice that after dropping levels, 'Reformed Criminals' doesn't exist anymore.    
```{r message = F}
library(dplyr)
#Remove align level
comics <- comics %>%
  filter(align != 'Reformed Criminals') %>%
  droplevels()
table(comics$align, comics$gender)
```
**Graph: Side-by-side Barcharts**   

While a contingency table represents the counts numerically, it's often more useful to represent them graphically.  

* Reviews: **[ggplot()](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.2)** for **Data Visualization**   
```{r message = F}
library(ggplot2)
ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(position = "dodge") + #Decide whether to stack the graph or not.
  theme(axis.text.x = element_text(angle = 0 ))#Decide the element text of x axis displaying angle.
```
Besides counts graphs, we can also make proportional graphs.    
```{r}
tab <- table(comics$align, comics$gender)
options(scipen = 999, digits = 3) # Print fewer digits
prop.table(tab)     #Joint proportions
prop.table(tab, 2)  #'2' means Conditional on columns, '1' means Condintional on rows.
```
<h2 id = '2'></h2>
# **Numerical Data**    

<h2 id = '3'></h2>
# **Numerical Summaries**    

<h2 id = '4'></h2>
# **Case Study**