---
title: "Exploratory Data Analysis"
author: "Bourbon0212"
date: "2018/7/13"
output: html_document
---
<h2 id = '0'></h2>
* [Categorical Data](#1)    
* [Numerical Data](#2)    
* [Numerical Summaries](#3)    
* [Case Study](#4)

<h2 id = '1'></h2>
# **Categorical Data**    

**Contingency Table**   

* Reviews: **[Factors](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#1.5)**   
* Reviews: **[Dataframe](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#1.6)**   
* levels(): To show the levels of the factors.   
* Use '$' to select a specific row or column.   
* table(): To generate 'Contingency table'.   
```{r message = F}
library(readr)
comics <- read_csv('https://assets.datacamp.com/production/course_1796/datasets/comics.csv')
comics[1,]
levels(comics$align)
table(comics$align, comics$gender)
```
**Dropping Levels**   

The contingency table from the last exercise revealed that there are some levels that have very low counts.   

* Reviews: **[filter()](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.1)** for **Data Wrangling**    
* droplevels(): Drops levels that don't exist in a dataset.   

Notice that after dropping levels, 'Reformed Criminals' doesn't exist anymore.    
```{r message = F}
library(dplyr)
#Remove align level
comics <- comics %>%
  filter(align != 'Reformed Criminals') %>%
  droplevels()
table(comics$align, comics$gender)
```
**Graph: Side-by-side Barcharts**   

While a contingency table represents the counts numerically, it's often more useful to represent them graphically.  

* Reviews: **[ggplot()](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.2)** for **Data Visualization**   
```{r message = F}
library(ggplot2)
ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(position = "dodge") + #Decide whether to stack the graph or not.
  theme(axis.text.x = element_text(angle = 0 ))#Decide the element text of x axis displaying angle.
```   

Besides counts graphs, we can also make proportional graphs.        
```{r}
tab <- table(comics$align, comics$gender)
options(scipen = 999, digits = 3) # Print fewer digits
prop.table(tab)     #Joint proportions
prop.table(tab, 2)  #'2' means Conditional on columns, '1' means Condintional on rows.
```
By adding position = "fill" to geom_bar(), you are saying you want the bars to fill the entire height of the plotting window, thus displaying proportions and not raw counts.   
```{r}
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = 'fill') +
  ylab("proportion")#Rename y axis.
```

**Distribution of One Categorical Variable**    
Notice the difference between geom_bar() & geom_col(). geom_bar() is **ONE** variable graph. However geom_col is **TWO** variable graph.    
* Use facet_wrap(~*) to breakdown a barchart.
```{r}
#Change the order of the levels in align
comics$align <- factor(comics$align, 
                       levels = c('Bad', 'Neutral', 'Good'))
#Create plot of align
ggplot(comics, aes(x = align)) + 
  geom_bar()
#Plot of alignment broken down by gender
ggplot(comics, aes(x = align)) + 
  geom_bar() +
  facet_wrap(~ gender)
```
<h2 id = '2'></h2>
# **Numerical Data**    

**Faceted Histogram**
```{r message = FALSE}
library(readr)
cars <- read_csv('https://assets.datacamp.com/production/course_1796/datasets/cars04.csv')
cars[1,]
```
```{r warning = FALSE}
#Filter cars with 4, 6, 8 cylinders
common_cyl <- filter(cars, ncyl %in% c(4, 6, 8))
#Create box plots of city mpg by ncyl
ggplot(common_cyl, aes(x = as.factor(ncyl), y = city_mpg)) +
  geom_boxplot()
#Create overlaid density plots for same data
ggplot(common_cyl, aes(x = city_mpg, fill = as.factor(ncyl))) +
  geom_density(alpha = .3)
```

According to this graph, we can conclude that:    

+ The highest mileage cars have 4 cylinders.    
+ The typical 4 cylinder car gets better mileage than the typical 6 cylinder car, which gets better mileage than the typical 8 cylinder car.    
+ Most of the 4 cylinder cars get better mileage than even the most efficient 8 cylinder cars.    
+ The variability in mileage of 8 cylinder cars seem much smaller than that of 4 cylinder cars.   

**Distribution of One Numerical Variable**    
```{r warning = FALSE, message = FALSE}
#Create hist of horsepwr
cars %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram() +
  ggtitle('horsepower')
#Create hist of horsepwr for affordable cars
cars %>% 
  filter(msrp < 25000) %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram() +
  xlim(c(90, 550)) +
  ggtitle('horsepower')
```

Compare these two graphs, notice that afordable cars tend to has less horesepower. NO cars in the less expensive range exceed 250 horsepower.

**BinWidth**    
As binwidth becomes larger, the graphh becomes smoother. It's crucial to derermine how large is the binwidth.
```{r warning = FALSE, message = FALSE}
#Create hist of horsepwr with binwidth of 3
cars %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram(binwidth = 3) +
  ggtitle('Horsepower binwidth = 3')
#Change binwidth to 30
cars %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram(binwidth = 30) +
  ggtitle('Horsepower binwidth = 30')
#Change binwidth to 60
cars %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram(binwidth = 60) +
  ggtitle('Horsepower binwidth = 60')
```

**Density & BoxPlots**    

* Reviews: **[Boxplots](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.4)**    
Both density plots and box plots display the central tendency and spread of the data, but the box plot is more robust to **outliers**.
```{r warning = FALSE, message = FALSE}
#Exclude outliers from data
cars_no_out <- cars %>%
  filter(msrp < 100000)
#Construct box plot of msrp using the reduced dataset
cars_no_out %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
#Create plot of city_mpg
cars %>% 
  ggplot(aes(x = 1, y = city_mpg)) +
  geom_boxplot()
#Create plot of width
cars %>% 
  ggplot(aes(x = width)) +
  geom_density()
```

Because the city_mpg variable has a much wider range with its outliers, it's best to display its distribution as a box plot.     

**Higher Dimensions**
```{r warning = FALSE, message = FALSE}
common_cyl %>%
  ggplot(aes(x = hwy_mpg)) +
  geom_histogram() +
  facet_grid(ncyl ~ suv) +
  ggtitle('ncyl&suv')
```

Across both SUVs and non-SUVs, mileage tends to decrease as the number of cylinders increases according to the picture.   

<h2 id = '3'></h2>
# **Numerical Summaries**    

**Center Measures: Mean, Median &Mode**     

* Reviews: **[Grouping and Summarizing](https://bourbon0212.github.io/NTU-CS-X/Week1/hw1.html#3.3)**    

```{r}
library(gapminder)
#Compute groupwise mean and median lifeExp
gap2007 <- filter(gapminder, year == 2007)
gap2007 %>%
  group_by(continent) %>%
  summarize(mean(lifeExp),
            median(lifeExp))
#Generate box plots of lifeExp for each continent
gap2007 %>%
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```

**Variance Measures: sd, var,diff & IQR**   

* sd(): Calculate **Standard Deviation**.   
* var(): Calculate **Variance** which equals to sd squared. ```sum((x - mean(x))^2)/(n-1)```    
* diff(range()): Calculate **Range**    
* IQR(): Calculate **Interquartile Range**    
Like mean and standard deviation, median and IQR measure the central tendency and spread, respectively, but are robust to outliers and non-normal data.   
```{r}
#Compute groupwise measures of spread
gap2007 %>%
  group_by(continent) %>%
  summarize(sd(lifeExp),
            IQR(lifeExp),
            n())
#Generate overlaid density plots
gap2007 %>%
  ggplot(aes(x = lifeExp, fill = continent)) +
  geom_density(alpha = 0.3)#alpha is about the transparency.
```

**Shapes & Transformation**
```{r}
#Create density plot of population
gap2007 %>%
  ggplot(aes(x = pop)) +
  geom_density()
#Create density plot of log(pop)
gap2007 %>%
  ggplot(aes(x = log(pop))) +
  geom_density()
```

The difference is much easier to see on a log scale.    

**Outliers**    
```{r}
#Outliers
gap_asia <- gap2007 %>%
  filter(continent == 'Asia') %>%
  mutate(is_outlier = lifeExp < 50)
#Create box plot of lifeExp with outliers
gap_asia %>%
  ggplot(aes(x = 1, y = lifeExp)) +
  geom_boxplot()
#Create box plot of lifeExp without outliers
gap_asia %>%
  filter(!is_outlier) %>%
  ggplot(aes(x = 1, y = lifeExp)) +
  geom_boxplot()
```

<h2 id = '4'></h2>
# **Case Study**

**Spam & Num_char**
```{r  warning = FALSE, message = FALSE}
library(openintro)
head(email, 3)
#Compute summary statistics
email %>%
  group_by(spam) %>%
  summarize(median(num_char), IQR(num_char))
#Create plot
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = spam, y = log_num_char, group = spam)) +
  geom_boxplot()
```

According to the graphs, the median length of not-spam emails is greater than that of spam emails.    

**Spam & exclaim_mess**

```{r  warning = FALSE, message = FALSE}
#Compute center and spread for exclaim_mess by spam
email %>%
  group_by(spam) %>%
  summarize(median(exclaim_mess), IQR(exclaim_mess))
#Create plot for spam and exclaim_mess
email %>%
  mutate(log_exclaim_mess = log(exclaim_mess)) %>%
  ggplot(aes(x = log_exclaim_mess)) +
  geom_histogram() + 
  facet_wrap(~ spam)
#Create plot of proportion of spam by image
email %>%
  mutate(has_image = image > 0) %>%
  ggplot(aes(x = has_image, fill = spam, group = spam)) +
  geom_bar(position = 'fill')
```

Obviously, an email without an image is more likely to be not-spam than spam.   

**Strategy for Testing**
```{r}
#Test if images count as attachments
sum(email$image > email$attach)
```

Since image is never greater than attach, we can infer that images are counted as attachments.    

**Question 1: **    
For emails containing the word "dollar", does the typical spam email contain a greater number of occurrences of the word than the typical non-spam email?   
Create a summary statistic that answers this question.    
```{r}
email %>%
  filter(dollar > 0) %>%
  group_by(spam) %>%
  summarize(median(dollar))
```

**Question 2: **    
If you encounter an email with greater than 10 occurrences of the word "dollar", is it more likely to be spam or not-spam?    
Create a barchart that answers this question.   
```{r}
email %>%
  filter(dollar > 10) %>%
  ggplot(aes(x = spam)) +
  geom_bar()
```

**Spam & NUmbers**
```{r}
# Reorder levels
email$number = factor(email$number, order = T, levels = c('none','small','big'))

# Construct plot of number
ggplot(email, aes(x = number)) +
  geom_bar() +
  facet_wrap(~spam)
```