---
title: "Happiness & Open Data (II)"
author: "Bourbon0212"
date: "2018年7月25日"
output: html_document
---

* [Happiness & Open Data (I)](https://bourbon0212.github.io/NTU-CS-X/Week3/Happiness_and_Open_Data.html)
* Preparation   
* Analysis
* Conclusion

# **Preparation**
```{r eval = F}
# Export the data we've collected from (I) to a csv
write.table(data_table, file = "data.CSV", sep = ",")
```
```{r message = F}
# Import data
library(formattable)
library(dplyr)
data <- read.csv('./data.csv') 
data %>%
  formattable(align = "l") %>%
  head(5)
```

<br>

# **Analysis**    

## **Openness & Region**   

### Errorbar Plot
```{r message = F}
library(ggplot2)
library(Hmisc)
data <- filter(data, !is.na(Region))
data %>%
  ggplot(aes(x = Region, y = Openness)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    # fun.args = list(mult = 1) to have a errorbar that spans over one standard deviation in both directions.
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'Openness', title = 'Openness & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$Openness), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(Openness ~ Region, data = data))
```

According to this plot & anova test, we can conclude that:     

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.     
* In `Sub-Saharan Africa` and `Central and Easttern Europe` the openness of some countries are `0`.     
* It's obvious that `Region` has noting to do with `Openness` in Ocenia.    

<br>

## **Happiness & Region**    

### Errorbar Plot
```{r}
data %>%
  ggplot(aes(x = Region, y = Happiness)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'Happiness', title = 'Happiness & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$Happiness), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(Happiness ~ Region, data = data))
```
According to this plot & anova test, we can conclude that:   

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.     
* Oceanian are the happiest people of all the regions.    
* The countries of `Happiness` in `SouthEastern Asia` & `Middle East & Northern Africa` are the least centralized.    

<br>

## **GDP & Region**  

### Errorbar Plot
```{r}
data %>%
  ggplot(aes(x = Region, y = GDP)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'GDP', title = 'GDP & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$GDP), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(GDP ~ Region, data = data))
```
According to this plot & anova test, we can conclude that:   

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.   
* The countries of `GDP` in `SouthEastern Asia` & `Middle East & Northern Africa` are the least centralized.
* The distribute of this plot is **similar** to the previous one.   
* So I make a prediction that the correlation between `Happiness` & `GDP` is high.    

### Regression
```{r}
m2 <- lm(GDP ~ Openness + Region - 1, 
         data = data)
coefplot(m2, xlab = 'GDP Estimated', ylab = 'Regress Varaibles', title = 'Response = GDP')
```

## **Happiness & GDP**

### Linear Model
Let's draw a plot that confirmed this assumption and create the linear model of this dataset.      
```{r}
data %>%
  ggplot(aes(x = Happiness, y = GDP)) +
  geom_jitter(shape = 16, alpha = 0.5) +
  stat_smooth(method = "loess", aes(col = "LOESS"), se = F) +
  stat_smooth(method = "lm", aes(col = "Linear"), se = F) +
  scale_color_discrete("Model") +
  theme(panel.grid = element_blank(), panel.background = element_blank())
```

```{r}
formattable(summarise(data, Observation = n(), Correlation = cor(GDP, Happiness)))
```
The correlation here is `0.7969` which suggests that my assumption is correct.    
Here are some details of this linear model.    
```{r}
summary(lm(GDP ~ Happiness, data = data))
```

We can also facet the plot according to different region.     
Except those regions with only one or two countries, the correlation between `GDP` & `Happiness` are positive.   
```{r}
ggplot(data, aes(x = Happiness, y = GDP, group = Region)) +
  geom_jitter(shape = 16, alpha = 0.5) +
  geom_smooth(method = 'lm', se = F) +
  facet_grid(.~Region) +
  theme(strip.text.x = element_text(angle = 90), strip.background = element_blank(),
        panel.grid = element_blank(), panel.background = element_blank())
```