---
title: "Happiness & Open Data (II)"
author: "Bourbon0212"
date: "2018年7月25日"
output: html_document
---

* [Happiness & Open Data (I)](https://bourbon0212.github.io/NTU-CS-X/Week3/Happiness_and_Open_Data.html)
* [Preparation](#1)   
* [Analysis](#2)    
    + [Openness & Region](#2.1)   
    + [Happiness & Region](#2.2)    
    + [GDP & Region](#2.3)    
    + [Happiness & GDP](#2.4)   
* [Conclusion](#3)    

<h2 id = '1'></h2>
# **Preparation**
```{r eval = F}
# Export the data we've collected from (I) to a csv
write.table(data_table, file = "data.CSV", sep = ",")
```
```{r message = F}
# Import data
library(formattable)
library(dplyr)
setwd("D:/GitHub/NTU-CS-X/Week4")
data <- read.csv('./data.csv') 
data %>%
  formattable(align = "l") %>%
  head(5)
```

<br>

<h2 id = '2'></h2>
# **Analysis**    

<h3 id = '2.1'></h3>
## **Openness & Region**   

### Errorbar Plot
```{r message = F}
library(ggplot2)
library(Hmisc)
data <- filter(data, !is.na(Region))
data %>%
  ggplot(aes(x = Region, y = Openness)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    # fun.args = list(mult = 1) to have a errorbar that spans over one standard deviation in both directions.
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'Openness', title = 'Openness & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$Openness), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(Openness ~ Region, data = data))
```

According to this plot & anova test, we can conclude that:     

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.     
* In `Sub-Saharan Africa` and `Central and Easttern Europe` the openness of some countries are `0`.     
* It's obvious that `Region` has noting to do with `Openness` in Ocenia.    

### Regression
```{r}
library(coefplot)
m1 <- lm(Openness ~ Region - 1, 
         data = data)
coefplot(m1, xlab = 'Openness Estimated', ylab = 'Regress Varaibles', title = 'Response = Openness')
```

```{r}
# 把資料與迴歸分析的預測值、殘差與影響度放進資料
fit_m1 <- data.frame(data[, c("Region","Openness")], fitted = fitted(m1), resid = resid(m1),
                     infl = influence(m1)$hat)
```

```{r}
# 依地區疊合真實觀測值與預測值
ggplot(data = fit_m1, aes(x = Openness, group = Region )) +
 stat_density(geom = 'path', position = 'identity') +
 stat_density(geom = 'path', position = 'identity', aes(x = fitted)) +
 geom_vline(xintercept = c(with(data, tapply(Openness,Region, mean))), linetype = 'dotted')+
 facet_grid(Region ~ .) +
 labs(x = 'Openness', y = 'Density')
```

```{r}
# 看殘差分配，依快樂指數，檢視常態與變異數同質假設
ggplot(data = fit_m1, aes(x = scale(resid)), group = Region ) +
 stat_density(geom = 'path', position = 'identity', aes(col = Region)) +
 scale_linetype_manual(values = 10:1) +
 guides(linetype = guide_legend(reverse = TRUE)) +
 labs(x = '標準化殘差', y = 'Density') +
 theme(legend.position = c(.15, .8))
```

```{r}
library(lattice)
# 看看殘差的 Q-Q 圖，依快樂指數。檢視常態假設
qqmath(~ scale(resid) | Region, data = fit_m1, type = c('p', 'g', 'r'),
       xlab = '常態位數', ylab = '標準化殘差', layout = c(2, 3),
       pch = '.', cex = 2)
```

```{r, message = F, warning = F}
library(MASS)
# 畫預測值與殘差的散佈圖，檢查線性與等分散假設
ggplot(data = fit_m1, aes(x = fitted, y = scale(resid), group = Region )) +
  geom_point(pch = 20, size = 1) +
  stat_smooth(method = 'rlm', se = F) +
  facet_grid(Region ~ .) +
  labs(x = 'Openness Estimated', y = '標準化殘差')

# 呈現影響值（影響估計結果過大的值）與標準化殘差
ggplot(data = fit_m1, aes(x = infl, y = scale(resid), group = Region)) +
 geom_text(aes(label = rownames(fit_m1)), cex = 2) +
 geom_hline(yintercept = 0, linetype = 'dotted') +
 facet_grid(Region ~ .) +
 labs(x = '影響值', y = '標準化殘差')
```

```{r, warning = F}
# 看看影響值
summary(influence(m1)$hat)
```

<br>

<h3 id = '2.2'></h3>
## **Happiness & Region**    

### Errorbar Plot
```{r}
data %>%
  ggplot(aes(x = Region, y = Happiness)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'Happiness', title = 'Happiness & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$Happiness), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(Happiness ~ Region, data = data))
```
According to this plot & anova test, we can conclude that:   

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.     
* Oceanian are the happiest people of all the regions.    
* The countries of `Happiness` in `SouthEastern Asia` & `Middle East & Northern Africa` are the least centralized.    


### Regression
```{r}
library(coefplot)
m2 <- lm(Happiness ~ Region - 1, 
         data = data)
coefplot(m2, xlab = 'Happiness Estimated', ylab = 'Regress Varaibles', title = 'Response = Happiness')
```

```{r}
# 把資料與迴歸分析的預測值、殘差與影響度放進資料
fit_m2 <- data.frame(data[, c("Region","Happiness")], fitted = fitted(m2), resid = resid(m2),
                     infl = influence(m2)$hat)
```

```{r}
# 依地區疊合真實觀測值與預測值
ggplot(data = fit_m2, aes(x = Happiness, group = Region )) +
 stat_density(geom = 'path', position = 'identity') +
 stat_density(geom = 'path', position = 'identity', aes(x = fitted)) +
 geom_vline(xintercept = c(with(data, tapply(Happiness,Region, mean))), linetype = 'dotted')+
 facet_grid(Region ~ .) +
 labs(x = 'Happiness', y = 'Density')
```

```{r}
# 看殘差分配，依快樂指數，檢視常態與變異數同質假設
ggplot(data = fit_m2, aes(x = scale(resid)), group = Region ) +
 stat_density(geom = 'path', position = 'identity', aes(col = Region)) +
 scale_linetype_manual(values = 10:1) +
 guides(linetype = guide_legend(reverse = TRUE)) +
 labs(x = '標準化殘差', y = 'Density') +
 theme(legend.position = c(.15, .8))
```

```{r}
library(lattice)
# 看看殘差的 Q-Q 圖，依快樂指數。檢視常態假設
qqmath(~ scale(resid) | Region, data = fit_m2, type = c('p', 'g', 'r'),
       xlab = '常態位數', ylab = '標準化殘差', layout = c(2, 3),
       pch = '.', cex = 2)
```

```{r, message = F, warning = F}
library(MASS)
# 畫預測值與殘差的散佈圖，檢查線性與等分散假設
ggplot(data = fit_m2, aes(x = fitted, y = scale(resid), group = Region )) +
  geom_point(pch = 20, size = 1) +
  stat_smooth(method = 'rlm', se = F) +
  facet_grid(Region ~ .) +
  labs(x = 'Happiness Estimated', y = '標準化殘差')

# 呈現影響值（影響估計結果過大的值）與標準化殘差
ggplot(data = fit_m2, aes(x = infl, y = scale(resid), group = Region)) +
 geom_text(aes(label = rownames(fit_m2)), cex = 2) +
 geom_hline(yintercept = 0, linetype = 'dotted') +
 facet_grid(Region ~ .) +
 labs(x = '影響值', y = '標準化殘差')
```

```{r, warning = F}
# 看看影響值
summary(influence(m2)$hat)
```

<br>

<h3 id = '2.3'></h3>
## **GDP & Region**  

### Errorbar Plot
```{r}
data %>%
  ggplot(aes(x = Region, y = GDP)) +
    geom_jitter(shape = 16, width = 0.1, alpha = 0.5) +
    stat_summary(geom = 'point', fun.y = mean, col = '#E41A1C') +
    stat_summary(geom = 'errorbar', fun.data = mean_sdl, fun.args = list(mult = 1), 
                 width = 0.3, col = '#E41A1C') +
    labs(y = 'GDP', title = 'GDP & Region - with T-tipped error bars') +
    theme(axis.title.y = element_blank(), panel.grid = element_blank(), panel.background = element_blank()) +
    geom_hline(yintercept = mean(data$GDP), linetype = 'dotted') +
    coord_flip()
```

### ANOVA
```{r}
anova(lm(GDP ~ Region, data = data))
```
According to this plot & anova test, we can conclude that:   

* Since `Pr(>F)` < `0.05`, the mean of each grop has significant difference.   
* The countries of `GDP` in `SouthEastern Asia` & `Middle East & Northern Africa` are the least centralized.
* The distribute of this plot is **similar** to the previous one.   
* So I make a prediction that the correlation between `Happiness` & `GDP` is high.    

### Regression
```{r}
library(coefplot)
m3 <- lm(GDP ~ Region - 1, 
         data = data)
coefplot(m3, xlab = 'GDP Estimated', ylab = 'Regress Varaibles', title = 'Response = GDP')
```

```{r}
# 把資料與迴歸分析的預測值、殘差與影響度放進資料
fit_m3 <- data.frame(data[, c("Region","GDP")], fitted = fitted(m3), resid = resid(m3),
                     infl = influence(m3)$hat)
```

```{r}
# 依地區疊合真實觀測值與預測值
ggplot(data = fit_m3, aes(x = GDP, group = Region )) +
 stat_density(geom = 'path', position = 'identity') +
 stat_density(geom = 'path', position = 'identity', aes(x = fitted)) +
 geom_vline(xintercept = c(with(data, tapply(GDP,Region, mean))), linetype = 'dotted')+
 facet_grid(Region ~ .) +
 labs(x = 'GDP', y = 'Density')
```

```{r}
# 看殘差分配，依快樂指數，檢視常態與變異數同質假設
ggplot(data = fit_m3, aes(x = scale(resid)), group = Region ) +
 stat_density(geom = 'path', position = 'identity', aes(col = Region)) +
 scale_linetype_manual(values = 10:1) +
 guides(linetype = guide_legend(reverse = TRUE)) +
 labs(x = '標準化殘差', y = 'Density') +
 theme(legend.position = c(.15, .8))
```

```{r}
library(lattice)
# 看看殘差的 Q-Q 圖，依快樂指數。檢視常態假設
qqmath(~ scale(resid) | Region, data = fit_m3, type = c('p', 'g', 'r'),
       xlab = '常態位數', ylab = '標準化殘差', layout = c(2, 3),
       pch = '.', cex = 2)
```

```{r, message = F, warning = F}
library(MASS)
# 畫預測值與殘差的散佈圖，檢查線性與等分散假設
ggplot(data = fit_m3, aes(x = fitted, y = scale(resid), group = Region )) +
  geom_point(pch = 20, size = 1) +
  stat_smooth(method = 'rlm', se = F) +
  facet_grid(Region ~ .) +
  labs(x = 'GDP Estimated', y = '標準化殘差')

# 呈現影響值（影響估計結果過大的值）與標準化殘差
ggplot(data = fit_m3, aes(x = infl, y = scale(resid), group = Region)) +
 geom_text(aes(label = rownames(fit_m3)), cex = 2) +
 geom_hline(yintercept = 0, linetype = 'dotted') +
 facet_grid(Region ~ .) +
 labs(x = '影響值', y = '標準化殘差')
```

```{r, warning = F}
# 看看影響值
summary(influence(m3)$hat)
```

<br>

<h3 id = '2.4'></h3>
## **Happiness & GDP**

### Linear Model
Let's draw a plot that confirmed this assumption and create the linear model of this dataset.      
```{r}
data %>%
  ggplot(aes(x = Happiness, y = GDP)) +
  geom_jitter(shape = 16, alpha = 0.5) +
  stat_smooth(method = "loess", aes(col = "LOESS"), se = F) +
  stat_smooth(method = "lm", aes(col = "Linear"), se = F) +
  scale_color_discrete("Model") +
  theme(panel.grid = element_blank(), panel.background = element_blank())
```

```{r}
formattable(summarise(data, Observation = n(), Correlation = cor(GDP, Happiness)))
```
The correlation here is `0.7969` which suggests that my assumption is correct.    
Here are some details of this linear model.    
```{r}
summary(lm(GDP ~ Happiness, data = data))
```

We can also facet the plot according to different region.     
Except those regions with only one or two countries, the correlation between `GDP` & `Happiness` are positive.   
```{r}
ggplot(data, aes(x = Happiness, y = GDP, group = Region)) +
  geom_jitter(shape = 16, alpha = 0.5) +
  geom_smooth(method = 'lm', se = F) +
  facet_grid(.~Region) +
  theme(strip.text.x = element_text(angle = 90), strip.background = element_blank(),
        panel.grid = element_blank(), panel.background = element_blank())
```

Now, we're going to analysis whether other factors have influence on `GDP` or not.
```{r, warning = F, message = F}
analysis<-data[,c("GDP","Happiness","Openness","Freedom","Trust","Health")]
# 看看基本統計量
colMeans(analysis)
# 兩兩散佈圖
library(heplots)
scatterplotMatrix(~ GDP + Happiness + Openness + Freedom + Trust + Health, data = analysis,
  pch = '.', cex = 3, smooth = FALSE, reg.line = FALSE, ellipse = TRUE,
  diagonal = 'none', lower.panel = NULL)
# 相關性
library(corrplot)
corrplot(cor(analysis), method = 'ellipse', order = 'hclust', addrect = 6,
         type = 'upper', tl.pos = 'tp')
corrplot(cor(analysis), add = TRUE, type = 'lower', method = 'number',
         order = 'hclust', col = 'black', diag = FALSE, tl.pos = 'n', cl.pos = 'n')
# 解釋5個變項
summary(m4 <- lm(GDP ~ Happiness + Openness + Freedom + Trust + Health, data = analysis))

# 相關性繪圖
library(effects)
plot(allEffects(m4), main = '', grid = T)
```

Apparently, we can see the correlation between GDP & Happiness, GDP & Health, GRP & Trust are relatively high.    
Here's the summary of these linear model.   
```{r}
library(lm.beta)
summary(lm.beta(m4))
```

<br>

<h2 id = '3'></h2>
# **Conclusion**

First, the tests tell us that the significance is strong between each region no matter what variables we choose in openness, happiness or GDP.       
Second, according to the scatter plots of 'GDP & Region', 'Happiness & Region',     
we conclude that there is a strong relationship between these two variables.    
Third, we made some plots about these two variables and confirmed the assumption.   
Finally, we add some new variables to the model and see their influnces on GDP.     